Notes and other text

### 2023-03-29 doc/manuscript/translocation.qmd
Minor edits and add several comments

Despite the catastrophic impact of Bd on MYL frogs in which most Bd-naive populations are driven to extirpation following Bd arrival [@vredenburg2010], some populations have persisted after epizootics [@briggs2010] and are now recovering [@knapp2016].
Frogs in these recovering populations show reduced susceptibility to Bd infection, consistent with selection for more resistant genotypes [@knapp2016].
Furthermore, exploratory efforts to reintroduce frogs collected from recovering populations to sites from which frogs were previously extirpated indicate that population establishment is possible despite ongoing Bd infection [@joseph2018].
Here, we report the results of (i) an analysis to evaluate genomic signatures of selection following MYL frog exposure to Bd, (ii) a greatly expanded MYL frog reintroduction effort, and (iii) population viability modeling to assess the long-term persistence of reintroduced populations.
Collectively, the results show that MYL frogs generally meet the four conditions for reintroduction feasibility outlined above, and provide a novel example of how the reintroduction of resistant frogs can allow the large-scale recovery of this disease-impacted amphibian.
These results may have broad implications for the many species worldwide that are threatened with extinction by emerging diseases.
<!--# Bree comment: All of this was written before we knew the genomics had something cool to say. But now that the genomics feel a bit more elevated, I think this paragraph should be more bold. For example, we might want more here or in the Discussion to make it clear how the results of each of these sections are super important on their own. And that combining them is really amazing. I think the current tone is appropriately factual, but at some point we need to be louder about the novelty involved and the incredible amount of work that this represents.  This version feels like, we did a,b,c and so MYLF meet the conditions for feasibility (which makes it sound like we've only just gotten to the *starting* point of the study. I think we need to go further than saying they meet the four conditions of a specific conservation action. This is where beefing up the conclusions about evolutionary rescue seem appropriate. Also, genomic analyses typically come last in our papers as a dive into possible mechanistic explanations for observed patterns in the field. I think the order of the primary sections makes sense given the way you set up the argument about evolutionary rescue. That said, the order combined with the tone feels like it's contributing to the sense that the genomics are a box to be checked rather than contributing something exciting to the overall story. Now that Allie has added new analyses, I think we could make a slightly bolder set-up about how the parts fit together to make a surprisingly strong whole -->

@@

@@

<!--# Roland comment: 2 of the 3 donor populations are included in the selection study. -->

@@

To identify site and frog-specific drivers of post-translocation survival, we developed a multilevel model in a Bayesian framework.
Our specific interest was whether Bd load had an important effect on adult survival, as would be expected if the effect of frog resistance was insufficient to overcome the negative effect of Bd.
<!--# Allie comment: This phrasing is a bit confusing for me. --> The best model of 1-year frog survival identified several important predictors, but Bd load at the time of translocation was not among them (bd_load; @fig-survival-postdens SI).
Important predictors included winter severity in the year following translocation (snow_t1), site elevation (elevation), and donor population (donor_72996; @fig-cond-effects).
Males had somewhat higher survival than females, but this effect was only marginally important (sex_male; @fig-cond-effects, @fig-survival-postdens SI).
In addition to Bd load, effects of frog size (length), winter severity in the year of translocation (snow_t), day of year on which a translocation was conducted (day), and translocation order (order_first) were all unimportant (@fig-survival-postdens SI).
Collectively, these results show that Bd load is not a predictor of the survival of translocated frogs, consistent with frogs in recovering populations having evolved sufficient resistance to minimize the adverse effects of Bd infection.
<!--# Allie comment: Maybe tolerance is a better word here? Since they were infected. Or maybe we need to explicitly define what we mean with resistance vs tolerance. Bree comment: I noticed this also – in the Intro, you do say that you are lumping both resistance and tolerance into the phrase "resistance" but you are talking about multiple things (being able to avoid infection, being able to fight infection, being able to tolerate infection, etc). Maybe at some point we want to address specifically that the genomic changes we are observing could contribute to multiple facets of resistance / tolerance.  And that we are lumping those together in an "evolved resistance" bin. --> In addition, the moderate Bd loads that characterized frogs following translocation and relatively high survival of frogs in the majority of more-recently translocated cohorts suggests that resistant frogs from recovering populations can maintain the benefits of resistance in non-natal habitats.
Finally, in three locations where longer capture-mark-recapture (CMR) time series allowed us to assess the survival of adults recruited to the population (i.e., progeny of translocated individuals), naturally recruited adults had equivalent or higher survival probabilities than the originally translocated adults (@fig-compare_surv_probs SI).
Therefore, frog resistance in non-natal habitats is maintained across generations.

<!--# Mark W: Regarding the above paragraph, what evidence do we have of "tolerance"? Are Bd loads significantly reduced in populations compared to epizootic levels? Can we convincingly state that the likely mechanism is more consistent with resistance or tolerance?  I ask because it would be nice to just use one of these terms in the paper and not get caught up in any resistance vs. tolerance discussions with reviewers.  Roland: Have a look at this paragraph of Knapp et al. PeerJ 2022: https://peerj.com/articles/12712/#p-61. We argue that there is evidence for resistance and tolerance. -->

>>>>

<!--# Roland will address some of Bree's points here -->Despite the catastrophic impact of Bd on MYL frogs in which most Bd-naive populations are driven to extirpation following Bd arrival [@vredenburg2010], some populations have persisted after epizootics [@briggs2010] and are now recovering [@knapp2016].
Frogs in these recovering populations show reduced susceptibility to Bd infection, consistent with selection for more resistant genotypes [@knapp2016].
Furthermore, exploratory efforts to reintroduce frogs collected from recovering populations to sites from which frogs were previously extirpated indicate that population establishment is possible despite ongoing Bd infection [@joseph2018].
Here, we report the results of (i) an analysis to evaluate genomic signatures of selection following MYL frog exposure to Bd, (ii) a greatly expanded MYL frog reintroduction effort, and (iii) population viability modeling to assess the long-term persistence of reintroduced populations.
Collectively, the results show that MYL frogs generally meet the four conditions for reintroduction feasibility outlined above, and provide a novel example of how the reintroduction of resistant frogs can allow the large-scale recovery of this disease-impacted amphibian.
These results may have broad implications for the many species worldwide that are threatened with extinction by emerging diseases.
<!--# Bree comment: All of this was written before we knew the genomics had something cool to say. But now that the genomics feel a bit more elevated, I think this paragraph should be more bold. For example, we might want more here or in the Discussion to make it clear how the results of each of these sections are super important on their own. And that combining them is really amazing. I think the current tone is appropriately factual, but at some point we need to be louder about the novelty involved and the incredible amount of work that this represents.  This version feels like, we did a,b,c and so MYLF meet the conditions for feasibility (which makes it sound like we've only just gotten to the *starting* point of the study. I think we need to go further than saying they meet the four conditions of a specific conservation action. This is where beefing up the conclusions about evolutionary rescue seem appropriate. Also, genomic analyses typically come last in our papers as a dive into possible mechanistic explanations for observed patterns in the field. I think the order of the primary sections makes sense given the way you set up the argument about evolutionary rescue. That said, the order combined with the tone feels like it's contributing to the sense that the genomics are a box to be checked rather than contributing something exciting to the overall story. Now that Allie has added new analyses, I think we could make a slightly bolder set-up about how the parts fit together to make a surprisingly strong whole -->

@@

<!--# Need summary of importance of genetic results, also need better integration of different selection  results -->

@@

<!--# Roland comment: 2 of the 3 donor populations are included in the selection study. Add 2 sentences that summarize selection results and bridge to translocation section. -->

@@

To identify site and frog-specific drivers of post-translocation survival, we developed a multilevel model in a Bayesian framework.
Our specific interest was whether Bd load had a negative effect on adult survival, as would be expected if frogs were highly susceptible to Bd infection.
<!--# Allie comment: This phrasing is a bit confusing for me. --> The best model of 1-year frog survival identified several important predictors, but Bd load at the time of translocation was not among them (bd_load; @fig-survival-postdens SI).
Important predictors included winter severity in the year following translocation (snow_t1), site elevation (elevation), and donor population (donor_72996; @fig-cond-effects).
Males had somewhat higher survival than females, but this effect was only marginally important (sex_male; @fig-cond-effects, @fig-survival-postdens SI).
In addition to Bd load, effects of frog size (length), winter severity in the year of translocation (snow_t), day of year on which a translocation was conducted (day), and translocation order (order_first) were all unimportant (@fig-survival-postdens SI).
Collectively, these results show that Bd load is not a predictor of the survival of translocated frogs, consistent with frogs in recovering populations having evolved sufficient resistance to minimize the adverse effects of Bd infection.
<!--# Allie comment: Maybe tolerance is a better word here? Since they were infected. Or maybe we need to explicitly define what we mean with resistance vs tolerance. Bree comment: I noticed this also – in the Intro, you do say that you are lumping both resistance and tolerance into the phrase "resistance" but you are talking about multiple things (being able to avoid infection, being able to fight infection, being able to tolerate infection, etc). Maybe at some point we want to address specifically that the genomic changes we are observing could contribute to multiple facets of resistance / tolerance.  And that we are lumping those together in an "evolved resistance" bin. --> In addition, the moderate Bd loads that characterized frogs following translocation and relatively high survival of frogs in the majority of more-recently translocated cohorts suggests that resistant frogs from recovering populations can maintain the benefits of resistance in non-natal habitats.
Finally, in three locations where longer capture-mark-recapture (CMR) time series allowed us to assess the survival of adults recruited to the population (i.e., progeny of translocated individuals), naturally recruited adults had equivalent or higher survival probabilities than the originally translocated adults (@fig-compare_surv_probs SI).
Therefore, frog resistance in non-natal habitats is maintained across generations.
